---
title: "Network Analysis - Tidyverse"
author: "`r Sys.getenv('USER')`"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message=FALSE,
                      warning=FALSE,
                      fig.height=3.5,
                      fig.width=6)

library(igraph)
library(ggraph)
library(visNetwork)
library(networkdata)
library(tidyverse)
library(lubridate)
library(here)
library(PerformanceAnalytics)
library(plotly)
library(scales)
library(gridExtra)
library(DT)

```

## Intro

Notes from DataCamp course ['Network Analytics with Tidyverse'](https://learn.datacamp.com/courses/network-analysis-in-the-tidyverse)

### Other References

* [Network Analytics with R blogpost](https://www.jessesadler.com/post/network-analysis-with-r/)

### Key Concepts 01

Networks are comprised of:

* **Nodes** (aka vertices, points): the 'things' in the network
    + must be identified by unique id
* **Edges** (aka links, ties): the connections between the things
    + can be directed or undirected relationships between the nodes
    + edge can have attributes
    + attributes can be used in weighting

Key packages in R for network analysis:

* igraph: network data manipulation and diagrams.
* ggraph: extends ggplot for network-relevant dimensions.
* visNetwork: interactive network diagrams.
* networkdata: ton of network datasets for exploration, demonstration, learning.

Where does tidyverse come in?

* Tidyverse concepts can be applied to network analysis and are sometimes integrated into other network analysis systems.

Where does igraph come in?

igraph is designed for organizing and working graph objects. Once a network is set up as igraph object, we can work with it in various ways, and this course is mainly about using igraph objects with tidyverse. 

### igraph data/setup

The course uses data on terrorist network involved in Madrid train bombing. I was able to obtain network data on the same group via networkdata package - it is identified there as 'covert_30'. [detail doc here](https://www.rdocumentation.org/packages/networkdata/versions/0.1.4/topics/covert_30)

```{r}
## madrid train bombing terrorist network data
madrid <- covert_30
## list of 25 igraph objects, based on different characteristics of the network
## can get info on all the igraph objects by subsetting the list:

#madrid[1:25] ## to look at them all

## or individual ones by:
madrid[[25]]
```

The top section provides info on the igraph object:

* IGRAPH: indicates igraph object ;)
* 03db05c: not sure, don't think it matters
* UNW-: indicates this is 'Undirected', 'Named', Weighted', '-' not Bipartite
* MADRID_TIE_YEAR: name of the graph
* attr: attributes, dimensions, and then showing edges based on vertex names.

Can save the object with whatever name and can also:

* get list of vertices
* get list of edges
* plot the network graph (simple plot - more to come on this later...)

```{r}
## SAVE selected object:
madrid_25 <- madrid[[25]]
## list of vertices (V) and edges (E)
V(madrid_25)
E(madrid_25)

## plot the network
plot.igraph(madrid_25)
```

### Tidyverse setup

#### Data frame from igraph object

For purposes of the course and working with tidyverse, we convert igraph object to data frame:

* can create data frames based on vertices, edges, or both at the same time (still two data frames)
* course uses separate vertices (nodes) and edges (ties) data frames

```{r}
## convert to data frame based on what= 'vertices', 'edges', 'both'
madrid_25_dfs <- get.data.frame(madrid_25, what='both')
madrid_25_dfv <- get.data.frame(madrid_25, what='vertices')
madrid_25_dfe <- get.data.frame(madrid_25, what='edges')

## for course examples:
g <- madrid_25
nodes <- madrid_25_dfv
ties <- madrid_25_dfe

head(nodes)
head(ties)
```

