---
title: "Network Analytics in R: Chapter 4 of 4"
author: "`r Sys.getenv('USER')`"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_depth: 4
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message=FALSE,
                      warning=FALSE,
                      fig.height=6,
                      fig.width=8)

library(igraph)
library(igraphdata) ## data sets for working with igraph
library(threejs) ## used for interactive plots at end

library(tidyverse)
library(lubridate)
library(here)
library(PerformanceAnalytics)
library(plotly)
library(scales)
library(gridExtra)
library(DT)

```

## Chapter 4: Identifying Special Relationships

Notes from DataCamp course ['Network Analysis in R'](https://learn.datacamp.com/courses/network-analysis-in-r) **Chapter 4: Identifying Special Relationships.**

## Key Concepts

* Assortativity: preferential attachwent of vertices to others that share similar attributes (numerical or categorical). Values from -1 to 1. 
* Reciprocity: proportion of edges that are both directions.
* Community detection: more advanced approach than assortativity and reciprocity for understanding groups/clusters/communities within a network
   + fastgreedy.community: like 'inside-out' method
   + edge.betweenness.community: sort of 'outside-in' method

## Get / Explore Data

Datacamp exercises use Forrest Gump movie character network data - pairs of actors who appear in at least one scene together. 

* For substitute, using **karate** club network data from **igraphdata** pkg.

```{r}
## get list of datasets in igraphdata
data(package='igraphdata')
## import data set
data('UKfaculty', package='igraphdata')

## get description of UKfaculty
UKfaculty
```

From this we can tell:

* Directed graph (D in D-W-)
* Not named graph: doesn't have 'name' attribute for vertices (no N in DNW)
* Weighted graph (W in D-W)
* 81 vertices
* 817 edges
* attributes: 
    + graph: Type (character), Date (character), Citation (character), Author (character) 
    + vertices: Group (number)
    + edges: weight (edges/number)

[Some explanation in official documentation here](https://igraph.org/r/doc/aaa-igraph-package.html)

## Preview Data

```{r}
plot(UKfaculty)
```


## Assortativity & Reciprocity

### Assortativity

Calculate assortativity:

* 'Group' is only vertex attribute, so will need to use that
* Group is already numeric, otherwise would have to convert to factor and then numeric 

```{r}
assortativity(UKfaculty, V(UKfaculty)$Group)
```

```{r}
assortativity.degree(UKfaculty, directed=TRUE)
```

### Randomization for Assessment

We now have values for assortativity and assortativity degree but hard to know how meaninful they are until we compare with randomization.

Randomization will help us to assess the values by comparing them to how likely the observed values are. 

```{r}
observed.assort <- assortativity(UKfaculty, V(UKfaculty)$Group)

results <- vector('list', 1000)
for(i in 1:10000){
  results[[i]] <- assortativity(UKfaculty, sample(V(UKfaculty)$Group))
}

## Plot
hist(unlist(results))
abline(v = observed.assort, col='red', lty=3, lwd=2)
```

Observed assortativity is not shown because it is beyond range of randomization histogram. So I guess it is relatively HIGH!

### Reciprocity

Check reciprocity

```{r}
reciprocity(UKfaculty)
```

## Community Detection

Back to Karate network data - same as used in DataCamp course.

```{r}
## import data set
data('karate', package='igraphdata')

## check data
karate
```

### fastgreedy.community

```{r}
fgc <- fastgreedy.community(karate)
sizes(fgc)
membership(fgc)
plot(fgc, karate)
```

### edge.betweenness.community

```{r}
ebc <- edge.betweenness.community(karate)
sizes(ebc)
plot(ebc, karate)
```

### Compare Plots

Plot fastgreedy.community and edge.betweenness.community side by side

```{r}
par(mfrow=c(1,2))
plot(fgc, karate)
plot(ebc, karate)

```

### leading.eigenvector.community

```{r}
lec <- leading.eigenvector.community(karate)
sizes(lec)
plot(lec, karate)
```

## Network visualizations

Other packages beyond **igraph** can be used:

* statnet
* ggnet
* ggnetwork
* ggraph
* visNetwork
* networkD3
* sigma
* rgexf (igraph to Gephi)
* **threejs**

### threejs for interactivity

Using threejs package for interactive network plots.

```{r}
UKfaculty <- set_vertex_attr(UKfaculty, 'color', value='dodgerblue')
graphjs(UKfaculty, vertex.size = 1)
```

Set size based on eigen centrality:

```{r}
g <- karate
ec <- as.numeric(eigen_centrality(g)$vector)

v <- 5 * sqrt(ec)

graphjs(g, vertex.size=ec)
```

Color based on community membership

```{r}
## create object containing memberships from fastgreedy.community detection above
i <- membership(fgc)
## check sizes
sizes(fgc)
## add color to each vertex based on community
karate <- set_vertex_attr(karate, "color", value=c("yellow","red","blue")[i])
## plot with threejs
graphjs(karate)

```

