---
title: "Network Analytics in R: Chapter 4 of 4"
author: "`r Sys.getenv('USER')`"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_depth: 4
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message=FALSE,
                      warning=FALSE,
                      fig.height=6,
                      fig.width=8)

library(igraph)
library(igraphdata) ## data sets for working with igraph

library(tidyverse)
library(lubridate)
library(here)
library(PerformanceAnalytics)
library(plotly)
library(scales)
library(gridExtra)
library(DT)

```

## Chapter 4: Identifying Special Relationships

Notes from DataCamp course ['Network Analysis in R'](https://learn.datacamp.com/courses/network-analysis-in-r) **Chapter 4: Identifying Special Relationships.**

### Key Concepts

* 

### Get / Explore Data

Datacamp exercises use Forrest Gump movie character network data - pairs of actors who appear in at least one scene together. 

* For substitute, using **karate** club network data from **igraphdata** pkg.

```{r}
## get list of datasets in igraphdata
#data(package='igraphdata')
## import data set
data('karate', package='igraphdata')

## get description of UKfaculty
karate
```

From this we can tell:

* Undirected graph (U in UNW)
* Named graph: has 'name' attribute for vertices (N in UNW)
* Weighted graph (W in UNW)
* not Bipartite (don't know what this is, not sure it comes up often)
* 34 vertices
* 78 edges
* attributes: 
    + graph: name (character), Citation (character), Author (character) 
    + vertices: Faction (number), name (character), label (character), color (number)
    + edges: weight (edges/number)

[Some explanation in official documentation here](https://igraph.org/r/doc/aaa-igraph-package.html)

To align with Forrest Gump data in Datacamp exercises we need to:

1. convert edges into data frame
2. drop the weighting to make it unweighted
3. rename cols to 'V1', 'V2'

```{r}
gdf <- igraph::as_data_frame(karate, what='edges')
gdf <- gdf %>% select(-weight) %>% rename(
  V1=from,
  V2=to
)
```

### Prep Data

Following along with exercises, convert data frame to graph

```{r}
g <- graph_from_data_frame(gdf, directed=FALSE)
```

### Heading


