---
title: "Network Analytics in R"
author: "`r Sys.getenv('USER')`"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message=FALSE,
                      warning=FALSE,
                      fig.height=3.5,
                      fig.width=6)

library(igraph)
library(igraphdata) ## data sets for working with igraph

library(tidyverse)
library(lubridate)
library(here)
library(PerformanceAnalytics)
library(plotly)
library(scales)
library(gridExtra)
library(DT)

```

## Intro 

Notes from DataCamp course ['Network Analysis in R'](https://learn.datacamp.com/courses/network-analysis-in-r).

Focus on social networks.

###>  Key Concepts 01

* Vertices: aka nodes, in this case, people in the social network.
* Edges: links between the vertices.
* Adjacency matrix: matrix with vertices as rows and columns, with values at the intersections representing edges
    + at minimum, 0 = no edge, 1 = edge exists between the vertices; can be higher values to represent attributes.
* Edgelist: list of combinations of vertices with edge values.

### Get Data

```{r}
## get list of datasets in igraphdata
data(package='igraphdata')
## import data set
data('UKfaculty', package='igraphdata')

## get description of UKfaculty
UKfaculty

```

From this we can tell:

* Directed graph
* not Named graph: no 'name' attribute for vertices
* Weighted graph
* not Bipartite (don't know what this is, not sure it comes up often)
* 81 vertices
* 817 edges
* attributes: 
    + graph: Type (character), Date (character), Citation (character), Author (character), 
    + vertices: Group (number)
    + edges: weight (edges/number)
* view of some edges

[Some explanation in official documentation here](https://igraph.org/r/doc/aaa-igraph-package.html)

#### Plot

Basic plot - we'll do better later. ;)

```{r}
## VERY useful for setting size - need to re-run for every code block with charts
par(mar=c(0,0,0,0))

## basic plot
plot(UKfaculty)
```

#### Additional Info

```{r}
## list of vertices
V(UKfaculty)
## list of edges
E(UKfaculty)

## number of edges
gsize(UKfaculty)
## number of vertices
gorder(UKfaculty)
```

#### Attributes

Can also get attributes of the network graph itself, vertices and edges:

```{r}
graph_attr(UKfaculty)
vertex_attr(UKfaculty)
edge_attr(UKfaculty)
```

### Add Attributes

* Note graph name change for convenience and to save original

Vertex attributes

```{r}
## using baby names data retrieved from https://www.ssa.gov/oact/babynames/limits.html, just for fun
## saved as: 'data/baby-names-1880'
new_attributes <- read_csv('data/baby-names-1880')
## get list same length as number of vertices
new_att <- sample_n(new_attributes, size=gorder(UKfaculty))

## change name for backup
ukf <- UKfaculty

ukf <- set_vertex_attr(ukf, name='leader', value=new_att$name)
ukf <- set_vertex_attr(ukf, name='type', value=new_att$gender)
ukf <- set_vertex_attr(ukf, name='size', value=new_att$count)

## check vertex attributes
vertex_attr(ukf)

## view the vertices
V(ukf)[[1:5]]
```

Sweet! More attributes to understand characteristics of the network.

Edge Attributes: set and view

```{r}
## add an attribute to edges to indicate strength
strength <- sample(x=seq(1:20), size=gsize(ukf), replace=TRUE)

ukf <- set_edge_attr(ukf, name='strength', value=strength)

edge_attr(ukf)
```

### Subset Edges

```{r}
## view first 5 in edge list
E(ukf)[[1:5]]

## select edges that include '12' in either 'tail' or 'head'
E(ukf)[[inc(12)]]
## select edges where strength =17
E(ukf)[[strength==17]]
```


## Alternate dataset 

Can investigate this one in parallel - might be closer case to the course, as well as practical relevance.

### Creating graph from data frames

```{r}

t_nodes <- data.frame(names=c('BL','CL','EB','JA','JY','IK','KK','MM','SN','TF','VN'),
                    role=c('Sr Analyst','Adv Analyst', 'Jr Analyst','Analyst','Sr Manager','Analyst','Engineer','Manager','Sr Analyst','Sr Analyst','Sr Analyst'))

t_edges <- tribble(
  ~from, ~to, ~weight,
  "BL", "JY", 4,
  "CL", "JY", 3,
  "EB", "MM", 2,
  "JA", "JY", 1,
  "IK", "MM", 1,
  "KK", "JY", 1,
  "MM", "JY", 1,
  "SN", "MM", 3,
  "TF", "JY", 1,
  "VN", "JY", 2)


t <- graph_from_data_frame(d=t_edges, directed=TRUE, vertices=t_nodes)

t
vertex_attr(t)
edge_attr(t)
```

#### Plot

```{r}
## VERY useful for setting size - need to re-run for every code block with charts
par(mar=c(0,0,0,0))
plot(t)
```

```{r}
## add attribute to network vertices
t <- set_vertex_attr(t, name='gender', value=c('M','F','F','M','M','F','M','F','M','M','F'))
vertex_attr(t)
```

```{r}
## set par for better sizing
par(mar=c(0,0,0,0))

V(t)$color <- ifelse(V(t)$gender=='F','orange','dodgerblue')

plot(t, vertex.label.color="white")
```

* node color based on gender
* labels set to white

### Other Plot Options

```{r}
## very useful in increasing default size! need to run with every code block that has plot
par(mar=c(0,0,0,0))

## recommended in other course
plot(t, layout=layout_with_kk(t))

## circle
plot(t, layout=layout_in_circle)

## fruchterman-reingold
plot(t, layout=layout_with_fr)

## layout as tree (doesn't seem to work - could be due to graph structure)
plot(t, layout=layout_as_tree(t))

## layout as star - looks like some inappropriate crossing (in some versions)
plot(t, layout=layout_as_star)

## igraph's choice
m1 <- layout_nicely(t)
plot(t, layout=m1)

```

#### Parameters

[Info on igraph plot parameters](https://www.r-graph-gallery.com/248-igraph-plotting-parameters.html)

* Node shape & size
* Edge weight
* Layout algo

```{r}
## set par for better sizing
par(mar=c(0,0,0,0))

## NODE shape & size
## shape options are basically circle or square :( (there are variations but hard to tell what difference)
V(t)$shape <- ifelse(V(t)$role=='Sr Manager','square',
                     ifelse(V(t)$role=='Manager', 'rectangle','circle'))
## default size = 15; max appears to be 24
V(t)$size <- ifelse(V(t)$role=='Sr Manager',22,
                     ifelse(V(t)$role=='Manager', 20,18))

## EDGE weight
## set edge width
w1 <- E(t)$weight

## LAYOUT
## set layout -> really, default 'layout_nicely' seems to be the best option, usually
lay <- layout_with_fr(t)

plot(t, vertex.label.color="white",
     edge.width=w1,
     edge.color='blue',
     layout=lay)
```

Edge weights may need more relative separation:

```{r}
table(t_edges$weight)
```

## Plotting Original Dataset

Explore plotting techniques with larger data set.

```{r, fig.height=8, fig.width=9}
par(mar=c(0,0,0,0))
#ukf
#vertex_attr(ukf)
#edge_attr(ukf)

## set node color
V(ukf)$color <- ifelse(V(ukf)$type=='F','red','dodgerblue')

## set size -> reduce variation to reasonable level
sizes <- V(ukf)$size
sizes <- ifelse(sizes>20, 20, ifelse(
  sizes<12,12, sizes
))

## plot with parameters
plot(ukf, edge.width=E(ukf)$weight,
     edge.color='blue',
     vertex.size=sizes,
     vertex.label.color='white',
     layout=layout_with_fr(ukf))
```

## Directional example

* check out another data set for consideration from igraphdata collection
* ideally looking for spread of contagious disease

```{r}
## possible alternative data set for exploration
## check handy options
#data(package='igraphdata')

## import prospective data set
data('enron', package='igraphdata')
## check meta data
enron
## confirm count of nodes
gorder(enron)
## confirm count of edges
gsize(enron)

## get random sample of nodes - obviously could plunk in some numbers -> getting extra practice with 'sample' in this context ;)
V(enron)[[sample(x=seq(1:gorder(enron)), size=4, replace=FALSE)]]
## get random sample  of edges
E(enron)[[sample(x=seq(1:gsize(enron)), size=4, replace=FALSE)]]

```

* could be interesting for another time - can go with 't' for now; can go with ukr, since it has direction.

### Back to ukf

```{r}
## is there any edge between nodes (returns weight if available; otherwise 0=no, 1=yes)
## how many edges from one node to another
ukf[76,42]
## how many edges the other way?
ukf[42,76]

## show all incidents of edges going out from node 76
incident(ukf, 76, mode='out')
## show all edges in or out of 76
incident(ukf, 76, mode='all')

```

### Node Relationships

###> Key Concepts 02

* Neighbours: nodes immediately adjacent to another node, by direct edge connection.
    + can specify mode= all/in/out
    + neighbors(graph, node id, mode=c('all','in','out'))
* Intersection: use to identify combinations of relaionships around nodes
    + intersection(neighbor set 1, neighbor set 2)
* Distance: identify distance between nodes
    + Diagonal: max distance between nodes within network

#### In-and-Out

```{r}
## nodes that come IN to 76 (9 in this case)
neighbors(ukf, 76, mode='in')
## nodes that 76 goes OUT to (4 in this case)
neighbors(ukf, 76, mode='out')
## all nodes INTO or OUT FROM 76 (9+4=13 total)
neighbors(ukf, 76, mode='all')
```

(the 'from' here just refers to the id of the graph)

#### Neighbors & Intersections

```{r}
## nodes that both come out from a specified node and go in to another specified node 
n1 <- neighbors(ukf, 10, mode='out')
n2 <- neighbors(ukf, 76, mode='in')
i1 <- intersection(n1, n2)

n1
n2
i1
```

* These are the nodes can you trace back out of node 76 and back through node 10
    + they come out of 10 and into 76
    
#### Distance

```{r}
## which nodes are farthest apart?
farthest_vertices(ukf)

## matrix of distances
#distances(ukf)

## path sequence between nodes identified in farthest_vertices
get_diameter(ukf)

## identify nodes that reach selected node within 3 outbound connections
ego(ukf, order=3, nodes=72 , mode='in')

## all the nodes that are within 3 connections to selected node, either inbound or outbound
##  (includes all the PREVIOUS plus those reachable out from selected node)
ego(ukf, order=3, nodes=72, mode='all')
```

